---
title: "Transcriptomic classification using mixture of normal distributions"
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Transcriptomic classification using mixture of normal distributions}
  %\VignettePackage{splitTypeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
BiocStyle::markdown()
library(knitr)
```

<br>
**Package**: `r Rpackage("splitTypeR")`<br />
**Authors**: `r packageDescription("splitTypeR")[["Author"]]`<br />
**Version**: `r packageDescription("splitTypeR")$Version`<br />
**Compiled date**: `r Sys.Date()`<br />
**License**: `r packageDescription("splitTypeR")[["License"]]`<br />

# Licensing and citing

This package and the underlying `r Rpackage("splitTypeR")` code are 
distributed under the Artistic license 2.0. You are free to use and 
redistribute this software.

If you use this package for a publication, we would ask you to cite the 
following:

>  Deschênes A, Belleau P (2026). _splitTypeR: Transcriptomic classification  
  using mixture of normal distributions_. R package version 0.99.0,
  <https://github.com/adeschen/splitTypeR>.

# Introduction

`r Rpackage("splitTypeR")` can ...


# Loading `r Rpackage("splitTypeR")` package

As with any R package, the `r Rpackage("splitTypeR")` package should first be 
loaded with the following command:

```{r loadingPackage, warning=FALSE, message=FALSE}
library(splitTypeR)
```


# Description

TODO

# Session info

Here is the output of `sessionInfo()` on the system on which this document was 
compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

