---
title: "Transcriptomic classification using mixture of normal distributions"
output:
  BiocStyle::html_document:
    toc: true
bibliography: splitTypeR.bibtex
vignette: >
  %\VignetteIndexEntry{Transcriptomic classification using mixture of normal distributions}
  %\VignettePackage{splitTypeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
BiocStyle::markdown()
library(knitr)
```

<br>
**Package**: `r Rpackage("splitTypeR")`<br />
**Authors**: `r packageDescription("splitTypeR")[["Author"]]`<br />
**Version**: `r packageDescription("splitTypeR")$Version`<br />
**Compiled date**: `r Sys.Date()`<br />
**License**: `r packageDescription("splitTypeR")[["License"]]`<br />

# Licensing and citing

This package and the underlying `r Rpackage("splitTypeR")` code are 
distributed under the Artistic license 2.0. You are free to use and 
redistribute this software.

If you use this package for a publication, we would ask you to cite the 
following:

>  Deschênes A, Belleau P (2026). _splitTypeR: Transcriptomic classification  
  using mixture of normal distributions_. R package version 0.99.0,
  <https://github.com/adeschen/splitTypeR>.

<br />

# Introduction

Cancer classification, using RNA sequencing, is now part of modern precision 
oncology. For some cancers, RNA classification provides deep insights into 
tumor aggressiveness and growth rates. For example, breast cancer subtypes 
(like Luminal-A, Luminal-B, HER2-enriched, or Basal-like) help forecast 
patient outcomes [@Cascianelli2020]. The Luminal-A is the most 
common molecular subtype of breast cancer, making up to 60% of all breast 
cancer cases. It also is associated with good prognosis. While 
Luminal-B tumors have a more aggressive phenotype and a worse 
prognosis [@Yersal2014].

Once gene signatures for RNA subtyping become widly accepted by major 
oncology organizations, they start to classify cancer subtypes, predict 
patient survival, and guide treatment.

From a research aspect, the classification of new samples using their 
transcriptional profiles can be done using different tools. For some cancers, 
machine learning models and dedicated software are available such as 
the Bioconductor Genefu package for breast cancer molecular 
subtyping [@Gendoo2016], and the CMScaller package for subtyping of 
colorectal cancer [@Eide2017].


The `r Rpackage("splitTypeR")` can ...

<br />

# Loading `r Rpackage("splitTypeR")` package

As with any R package, the `r Rpackage("splitTypeR")` package should first be 
loaded with the following command:

```{r loadingPackage, warning=FALSE, message=FALSE}
library(splitTypeR)
```


# Description

TODO

# Session info

Here is the output of `sessionInfo()` on the system on which this document was 
compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

