---
title: "Transcriptomic classification using mixture of normal distributions"
output:
  BiocStyle::html_document:
    toc: true
bibliography: splitTypeR.bibtex
vignette: >
  %\VignetteIndexEntry{Transcriptomic classification using mixture of normal distributions}
  %\VignettePackage{splitTypeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
BiocStyle::markdown()
suppressPackageStartupMessages({
  library(knitr)
})
```

<br />

**Package**: `r Rpackage("splitTypeR")`<br />
**Authors**: `r packageDescription("splitTypeR")[["Author"]]`<br />
**Version**: `r packageDescription("splitTypeR")$Version`<br />
**Compiled date**: `r Sys.Date()`<br />
**License**: `r packageDescription("splitTypeR")[["License"]]`<br />

# Licensing and citing

This package and the underlying `r Rpackage("splitTypeR")` code are 
distributed under the Artistic license 2.0. You are free to use and 
redistribute this software.

If you use this package for a publication, we would ask you to cite the 
following:

>  Deschênes A, Belleau P (2026). _splitTypeR: Transcriptomic classification  
  using mixture of normal distributions_. R package version 0.99.0,
  <https://github.com/adeschen/splitTypeR>.

<br />

# Introduction

Cancer classification, using RNA sequencing, is now part of modern precision 
oncology. For some cancers, RNA classification provides deep insights into 
tumor aggressiveness and growth rates. For example, breast cancer subtypes 
(like Luminal-A, Luminal-B, HER2-enriched, or Basal-like) help forecast 
patient outcomes [@Cascianelli2020]. The Luminal-A is the most 
common molecular subtype of breast cancer, making up to 60% of all breast 
cancer cases. It also is associated with good prognosis. While 
Luminal-B tumors have a more aggressive phenotype and a worse 
prognosis [@Yersal2014]. Once gene signatures for RNA subtyping become 
widly accepted by major oncology organizations, they start to classify cancer 
subtypes, predict patient survival, and guide treatment.

From a research aspect, the classification of new samples using their 
transcriptional profiles can be done using different tools. For some cancers, 
machine learning models and dedicated software are available such as 
the Bioconductor Genefu package for breast cancer molecular 
subtyping [@Gendoo2016], and the CMScaller package for subtyping of 
colorectal cancer [@Eide2017].

In the absence of dedicated software, unsupervised clustering with a 
heatmap visualization to classify transcriptomic samples with a 
gene signature is often employed. This method is relatively simple and 
offers an instant visual support. However, this method has a few drawbacks:

* _Lack of statistical metrics_: It doesn't 
provide a statistical probability score for a sample's class assignment; it 
mainly acts as a visual summary. 
* _Metric sensitivity_: Changing the clustering metric (ex: from "Euclidean" to 
"Pearson" distance) can alters the resulting clusters. 
* _Outlier distortion_: Extreme outliers can squash the color scale, and 
artificially create skewed subsets.

The __`r Rpackage("splitTypeR")`__ package resolves these issues by providing 
an automated statistical framework to classify heterogeneous biological 
samples based on gene signature lists, effectively isolating the 
signature-positive samples. 

This classification method is designed for bulk transcriptomic datasets.

<br />

# Installation

As with any R package, the `r Rpackage("splitTypeR")` package should first be 
loaded with the following command:

```{r loadingPackage, warning=FALSE, message=FALSE}
library(splitTypeR)
```

<br />

# Method

From a specific gene set, gene set enrichments are first calculated through the 
samples of a expression dataset with the Bioconductor GSVA 
package [@Hanzelmann2013].

Through an upscaling step that using a non-parametric permutation method, 
synthetic values are calculated to enable the calculation of a normal 
distribution using a mixture model approach implemented in the 
mixtools package [@JSSv032i06].   

# General workflow

TODO

The 

<br />

# Session info

Here is the output of `sessionInfo()` on the system on which this document was 
compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

